### 2025-04-30 conteudo/methods/methods.tex
add work 29/04

\todo[inline]{THE WHOLE SECTION NEEDS TO BE REMADE.}

@@

This section outlines our systematic approach for mitigating code duplication in the Linux kernel, details the methodology for selecting duplicated functions, and describes the ethnographic methodology for engaging with the Linux community.

\subsection{Systematic Approach to Mitigate Code Duplication}
\label{subsec:pipeline}

In C programming, communication between source files is achieved by creating header files that specify libraries \citep{Cbook}. Since the Linux kernel is primarily written in C, we eliminated code duplications by consolidating duplicated code into a single library, which would replace instances of duplication across the codebase.
%
For each duplicated function pair, we used the \textit{function information} functionality (described in Section \ref{subsec:functioncommand}) to locate all instances of the function in the code. This strategy resulted in a collection of duplicated functions and corresponding code files.

To identify shared code more effectively, we extended this approach to search for other common functions across all collection files. We then applied specific refactoring methods to each shared function. If the functions were identical across files, we removed the duplicates and created a single function in the library. If modifications existed, we applied case-specific refactoring.

Successful mitigations were submitted to the Linux community to gauge their feedback. If unsuccessful, we categorized the cases: some required specialist decisions beyond our expertise, and others were potentially non-refactorable. All mitigations were documented for research transparency at \url{https://github.com/LipArcanjo/master-artifacts}.

\subsection{Selection of Duplicated Functions}

As with the empirical analysis, evaluating every code duplication in the Linux kernel is impractical. 
Thus, we will limit the experiment to the AMD Display driver and used random sampling of duplicated functions
inferred by our tool to initiate the mitigation approach in Section \ref{subsec:pipeline}. 
We will choose a sample size of 30 duplicated functions, as studies indicate this size provides 
statistical significance \citep{sample1, sample2}.

For random selection, we will restrict the eligibility to function pairs with a similarity metric of 100\%. 
With this criterion, we randomly chose functions from eligible pairs uniformly.

\subsection{Ethnographic Methodology}

We conduct ethnographic research upon completing a code duplication mitigation following the systematic approach in Section \ref{subsec:pipeline}. We adopt a participant-observation methodology, contributing to the Linux kernel for the AMD Display driver and interacting with maintainers to incorporate our mitigations.

Linux kernel communication occurs via public mailing lists, where discussions are accessible to all. We collect data artifacts from the AMD Display driver mailing list, which records community feedback on our mitigation efforts, providing valuable insights into the proposed approach effectiveness.

Given our research timeline, we do not expect continuous interactions with the Linux community throughout all mitigations, as response time varies. Thus, we committed to applying the participant-observation approach for three to five successful mitigations and using these for our final discussion.

>>>>

@@

This section outlines our approach for mitigatin code duplication in the Linux kernel
and describes the ethnographic methodology for engagin with the Linux community. Firstly, 
we present the participant observation research method explored by the author. Then, 
we present the methodolody of the experiment done with the students.

\subsection{Participant observation method}

\subsection{Research with students}

\todo[inline]{Rsearch with students does not look good.}

In the Free Software Development course at University of São Paulo, ministred by the advisor 
of this work at the first semester of 2025, we have proposed to the students the possibility
of mitigating duplications in the IIO subsystem and the AMD display driver found with the Arkanjo tool.
The course had a total of X students, and of that, X students being undergraduate students
and Y begin graduate students.

On the course, the students are asked to form groups of one, two or three students and together 
make a contribution to a open source software. To auxiliate the students on this task, the 
professor and teaching assistants prepares multiple alternatives on how to contribute to open source projects, 
including simple contributions as updating documentation or code style fixes. In this context, 
we have proposed two alternatives for the students related to the tool.

For the first alternative, one of the teaching assistants used the Arkanjo tool in the IIO subsystem and
listed interesting duplications that the groups can approach, and it is up to the students on how
to refactor the code to mitigate the duplications and pass through the process to send patches to the IIO 
maintainers to review. For the second alternative, we propose a similar experience then the 
participant observation method, where the students pass through the whole process of executing the Arkanjo tool
in the AMD display driver, analyze the duplications found, create a patch to mitigate a duplication and sent
the patch to the AMD display driver's maintainers.

From the students enrolled in the in the course, X opted for the first alternative and 1 opted for the second
alternative, forming Y groups working on the first alternative and 1 group working on the second alternative. 
These number represent X\% of the students choosing for approching duplications them the other contribution 
alternatives, including more simpler and easy to be done.  It is worth highlighting that none of the students
that choose to mitigate duplications had made a contribution to the Linux kernel before,
meaning that the patch to mitigate the duplication being the first tentative to contribute to the kernel as
newcomers.

The groups have been request to document the experience and approaches to refactor the duplications and the
experience of submiting a patch to the Linux kernel in blogs. We had analyzed these blogs to undertand 
patterns of refactor used for mitigating the duplications, the experience on send the patches and the 
opinions of the subsystem maintainers on the mitigations. Appendix X list the blogs and patches made by the 
student.

For the teaching assistant and the student that has executed the Arkanjo tool to explore duplications, 
we have asked them to answer some questions related to the experience of using the tool and what could be
improved. The question asked can be found below and the answers can be found in Appendix Y.
